---
title: "Dutch Dashboard analyses:space"
author: "Dan Weinberger"
date: "1/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(reshape2)
library(rjags)
library(HDInterval)
library(ExcessILI)
library(lubridate)
library(dplyr)
library(zoo)
library(lme4)
library(readr)

source('./functions/refresh_data.R')
source('./functions/analyze_case_space.R')
source('./functions/analyze_nice_space.R')
source('./functions/delay_dist.R')

DownLoadLatest <- F
```

```{r}
pop1 <- read_delim('./Data/PopData/PopCBS.csv',delim=';')
pop1 <- pop1[,c('Regions',"PopulationOn1January_1")]
m1 <- readRDS('./Data/nice_data/2021_01_03_15_35.rds')
m1 <- m1[,c('Municipality_name','Municipality_code')]
m1 <- unique(m1)
pop1 <- merge(m1, pop1, all.x=T, by.x="Municipality_code", by.y="Regions")
names(pop1) <- c('code','muni','pop')
pop1$pop <- as.numeric(pop1$pop)
```

```{r}
case.data <- analyze_case_space_func(DownLoadLatest=DownLoadLatest)
```

```{r}
nice.data <- analyze_nice_space_func(DownLoadLatest=DownLoadLatest)
```

## Poisson random intercept
```{r}
nice.long <- nice.data$d1.c.long
nice.long <- merge(nice.long, pop1, by='muni')
nice.long$log.offset <- log(nice.long$days*nice.long$pop/100000)
nice.long$muni <- as.factor(nice.long$muni)
mod1 <- 
  glmer(hosp~ (period|muni ) +offset(log.offset), family='poisson', data=nice.long)
ran.effs <- ranef(mod1)[['muni']]
ran.effs$muni <- rownames(ran.effs)
names(ran.effs) <- c('int.nice','period.nice','muni')

```

Association between intercept and period effect
Higher incidence of hospitalization in period 1 associated with smaller increases in period 2
```{r}
plot( ran.effs$int.nice,ran.effs$period.nice)
abline(h=0)
```

```{r}
 ran.effs <- merge(ran.effs,pop1, by='muni')
```

Association of incidence in period 1 or increase with population
```{r}
plot(log(ran.effs$pop),ran.effs$int.nice )
plot(log(ran.effs$pop),ran.effs$period.nice)

```



### same for cases
```{r}
case.long <- case.data$d1.c.long
case.long <- merge(case.long, pop1, by='muni')

case.long$log.offset <- log(case.long$days*case.long$pop/100000)

case.long$muni <- as.factor(case.long$muni)
mod2 <- 
  glmer(cases~ (period|muni ) +offset(log.offset), family='poisson', data=case.long)
ran.effs.case <- ranef(mod2)[['muni']]
ran.effs.case$muni <- rownames(ran.effs.case)
names(ran.effs.case) <- c('int.case','period.case','muni')
```

```{r}
plot( ran.effs.case$int.case,ran.effs.case$period.case)
```

Compare hospitalization and case estimates. This shows general agreement that places that had less intense case reports also had less intense hospitalizations
```{r}
 ran.eff2 <- merge(ran.effs.case, ran.effs, by='muni')
plot(ran.eff2$period.nice, ran.eff2$period.case, bty='l', xlab='Log-Rate ratio NICE', ylab='Log-rate ratio positive tests')
abline(a=0, b=1)
```
Munis with lower rate in 2nd period vs first
```{r}
ran.eff2 <- ran.eff2[order(ran.eff2$period.nice),]
print(ran.eff2[ran.eff2$period.nice<0,])


```






