---
title: "Dutch Dashboard analyses:space"
author: "Dan Weinberger"
date: "1/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(reshape2)
library(rjags)
library(HDInterval)
library(ExcessILI)
library(lubridate)
library(dplyr)
library(zoo)
library(lme4)
library(readr)

source('./functions/refresh_data.R')
source('./functions/analyze_case_space.R')
source('./functions/analyze_nice_space.R')
source('./functions/delay_dist.R')

DownLoadLatest <- F
```

```{r}
pop1 <- read_delim('./Data/PopData/PopCBS.csv',delim=';')
pop1 <- pop1[,c('Regions',"PopulationOn1January_1")]
m1 <- readRDS('./Data/nice_data/2021_01_03_15_35.rds')
m1 <- m1[,c('Municipality_name','Municipality_code')]
m1 <- unique(m1)
pop1 <- merge(m1, pop1, all.x=T, by.x="Municipality_code", by.y="Regions")
names(pop1) <- c('code','muni','pop')
pop1$pop <- as.numeric(pop1$pop)
```

```{r}
case.data <- analyze_case_space_func(DownLoadLatest=DownLoadLatest)
```

```{r}
nice.data <- analyze_nice_space_func(DownLoadLatest=DownLoadLatest)
```

## Poisson random intercept
```{r}
nice.long <- nice.data$d1.c.long
nice.long <- merge(nice.long, pop1, by='muni')
nice.long$log.offset <- log(nice.long$days*nice.long$pop/100000)
nice.long$muni <- as.factor(nice.long$muni)
nice.long$period <- as.factor(nice.long$period)
mod1 <- 
  glmer(hosp~ (period|muni ) +offset(log.offset), family='poisson', data=nice.long)
ran.effs <- ranef(mod1)[['muni']]
ran.effs$muni <- rownames(ran.effs)
names(ran.effs) <- c('int.nice','period1.nice','period2.nice','muni')

```

Association between intercept and period effect
Higher incidence of hospitalization in period 1 associated with smaller increases in period 2
```{r, fig.width=5, fig.height=5}
plot( ran.effs$int.nice,ran.effs$period1.nice)
abline(h=0)

plot( ran.effs$int.nice,ran.effs$period2.nice)
abline(h=0)

plot( ran.effs$period1.nice,ran.effs$period2.nice)
abline(h=0)

#Per2/Per1 vs springbaseline
plot( ran.effs$int.nice,ran.effs$period2.nice-ran.effs$period1.nice, ylab='Increase in later winter vs Fall', xlab='Log(incidence Spring)')
abline(h=0)

#Increase in early fall vs spring vs vs spring baseline
plot( ran.effs$int.nice,ran.effs$period1.nice, ylab='Increase in fall vs Spring', xlab='Log(incidence Spring)')
abline(h=0)
```




```{r}
 ran.effs <- merge(ran.effs,pop1, by='muni')
```

Association of incidence in period 1 or increase with population
```{r}
plot(log(ran.effs$pop),ran.effs$int.nice )
plot(log(ran.effs$pop),ran.effs$period.nice)

```



### same for cases
```{r}
case.long <- case.data$d1.c.long
case.long <- merge(case.long, pop1, by='muni')

case.long$log.offset <- log(case.long$days*case.long$pop/100000)

case.long$muni <- as.factor(case.long$muni)
mod2 <- 
  glmer(cases~ (period|muni ) +offset(log.offset), family='poisson', data=case.long)
ran.effs.case <- ranef(mod2)[['muni']]
ran.effs.case$muni <- rownames(ran.effs.case)
names(ran.effs.case) <- c('int.case','period.case','muni')
```

```{r}
plot( ran.effs.case$int.case,ran.effs.case$period.case)
```

Compare hospitalization and case estimates. This shows general agreement that places that had less intense case reports also had less intense hospitalizations
```{r}
 ran.eff2 <- merge(ran.effs.case, ran.effs, by='muni')
# plot(ran.eff2$period1.nice, ran.eff2$period.case, bty='l', xlab='Log-Rate ratio NICE', ylab='Log-rate ratio positive tests')
# abline(a=0, b=1)
```

Munis with lower rate in 2nd period vs first
```{r}
ran.eff2 <- ran.eff2[order(ran.eff2$period1.nice),]
print(ran.eff2[ran.eff2$period1.nice<0,])


```

Simple plots
```{r}
f1 <- nice.data$d1.c
f1 <- merge(f1, pop1, by.x="Municipality_name" , by.y='muni')

f1$inc.per0 <- f1$'0'/f1$pop
f1$inc.per1 <- f1$'1'/f1$pop
f1$inc.per2 <- f1$'2'/f1$pop

f1$simp.rr1.2 <- (f1$'2'+0.5)/(f1$'1'+0.5)

plot(f1$inc.per0, f1$inc.per1)

plot(f1$inc.per1, f1$inc.per2)
plot(f1$inc.per0, f1$inc.per2)

symbols((f1$inc.per0+f1$inc.per1), f1$inc.per2, circles= log(f1$pop), inches=0.1, fg=NA, bg=rgb(0,0,0,0.1), bty='l')
abline(a=0, b=1)

symbols((f1$inc.per0+f1$inc.per1), log(f1$simp.rr1.2), circles= sqrt(log(f1$pop))/pi, inches=0.05, fg=NA, bg=rgb(0,0,0,0.1), bty='l')
abline(h=0)



```

Looks at changes over time by muni
```{r}
nice.wk <- nice.data$d1.c.wk
nice.dates <- nice.wk$Date_of_statistics


smooth.nice <- apply(nice.wk[,-c(1:2)], 2, stl.func, ts.date=nice.dates)

plot(smooth.nice[,'Utrecht'], type='l')
points(nice.wk[,'Utrecht'], type='l', col='gray')

nice.rr <- apply(smooth.nice, 2, rr.smooth.func)

check1 <- c("'s-Gravenhage" ,'Utrecht','Rotterdam', 'Amsterdam')
check2 <- c('Groningen','Haarlem','Amersfoort','Maastricht')

par(mfrow=c(1,2))
matplot(nice.dates,log(nice.rr[,check1]), type='l', ylim=log(c(0.5,5)))
abline(h=0)

matplot(nice.dates,log(nice.rr[,check2]), type='l', ylim=log(c(0.5,5)))
abline(h=0)

cum.cases <- apply(nice.wk[,-c(1:2)],2, cumsum)
common.munis <- sort(pop1$muni[pop1$muni %in% colnames(cum.cases)])
cum.cases <- cum.cases[,common.munis]
pop.sort <- pop1[order(pop1$muni),]
pop.sort <- pop.sort[!is.na(pop.sort$pop),]
pop.sort <- pop.sort[which(pop.sort$muni %in% colnames(cum.cases)),]

cum.inc <- sapply(colnames(cum.cases) , function(x) cum.inc =cum.cases[,x]/pop.sort$pop[pop.sort$muni==x]*100000) 

par(mfrow=c(1,2))
matplot(nice.dates,log(cum.inc[,check1]), type='l', ylim=(c(-2,6)))
abline(h=0)

matplot(nice.dates,log(cum.inc[,check2]), type='l', ylim=(c(-2,6)))
abline(h=0)



##Cum since sept 1

cum.cases <- apply(nice.wk[nice.dates >=as.Date('2020-09-01'),-c(1:2)],2, cumsum)
common.munis <- sort(pop1$muni[pop1$muni %in% colnames(cum.cases)])
cum.cases <- cum.cases[,common.munis]
pop.sort <- pop1[order(pop1$muni),]
pop.sort <- pop.sort[!is.na(pop.sort$pop),]

cum.inc <- sapply(colnames(cum.cases) , function(x) cum.inc =cum.cases[,x]/pop.sort$pop[pop.sort$muni==x]*100000) 

nice.dates.fall <- nice.dates[nice.dates >=as.Date('2020-09-01')]
par(mfrow=c(1,2))
matplot(nice.dates.fall,log(cum.inc[,check1]), type='l', ylim=(c(-2,6)))
abline(h=0)

matplot(nice.dates.fall,log(cum.inc[,check2]), type='l', ylim=(c(-2,6)))
abline(h=0)


```


```{r}
case.wk <- case.data$d1.c.wk
case.dates <- case.wk$Date_of_publication

#7 day move ave
smooth.case7 <- apply(case.wk[,-c(1:2)], 2, smooth.ave.func)
smooth.case.rr <- apply(smooth.case7, 2, rr.smooth.func)

#View(cbind.data.frame(case.dates,smooth.case7[,check1]))
```

```{r}
par(mfrow=c(1,2))
matplot(case.dates,smooth.case.rr[,check1], type='l',ylim=c(0.5,2.0))
abline(h=c(0.6,0.8,1.0), col='gray')

matplot(case.dates,smooth.case.rr[,check2], type='l',ylim=c(0.5,2.0))
abline(h=c(0.6,0.8,1.0), col='gray')
```
Same thing for NICE
```{r}
smooth.nice7 <- apply(nice.wk[,-c(1:2)], 2, smooth.ave.func)
smooth.nice.rr <- apply(smooth.nice7, 2, rr.smooth.func)
```

```{r}
matplot(nice.dates,smooth.nice.rr[,check1], type='l', ylim=c(0.5,3), bty='l')
abline(h=1,lty=2)

matplot(nice.dates,smooth.nice.rr[,check2], type='l', ylim=c(0.5,3), bty='l')
abline(h=1,lty=2)
```




Lansingerland: https://nltimes.nl/2021/01/22/large-scale-spread-b117-coronavirus-lansingerland-health-service
70K tested


```{r, fig.width=5, fig.height=6}
par(mfrow=c(2,1))
case.period <- which(case.dates>=as.Date('2020-09-01'))
plot(case.dates[case.period], case.wk$Lansingerland[case.period], type='l', main='Cases Lansingerland')
points(case.dates[case.period], smooth.case7[case.period,'Lansingerland'], type='l', col='red')
abline(v=as.Date('2021-01-14'), col='red',lty=2)

nice.period <- which(nice.dates>=as.Date('2020-09-01'))
plot(nice.dates[nice.period], nice.wk$Lansingerland[nice.period], type='l', main='Hos. Admissions Lansingerland')
points(nice.dates[nice.period], smooth.nice7[nice.period,'Lansingerland'], type='l', col='red')
abline(v=as.Date('2021-01-14'), col='red',lty=2)

```


